---
title: "Lab 06 - Sad plots"
author: "Elayna Seago"
date: "2/22/22"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

### Exercise 1

```{r}
staff <- read_csv("data/instructional-staff.csv")
```




```{r change-data-struc}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))

staff_long
```


```{r plot-bad-example}
staff_long %>%
  ggplot(aes(x = year, y = value, color = faculty_type)) +
  geom_line()

```

```{r fig.align="center", echo = FALSE,fig.width = 8}
staff_long %>%
  ggplot(aes(x = year, y = value, group = faculty_type, color = faculty_type)) +
  geom_line() +
  labs(title = "Instructional Employees From 1975 to 2011" , y = "Number in Position") 
  
```

```{r}
staff_long %>% 
  ggplot(aes(x = year, y = value , fill = faculty_type)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))+
  labs(title = "Proportion of Faculty in Each Position Over Time" , y = "Proportion")
  
```

### Exercises 2
```{r}
fisheries <- read_csv("data/fisheries.csv")
fisheries
```

```{r}
big_fisheries <- fisheries %>% 
  filter(total > 500000)

big_fisheries
```
### ### For my first attempted improvement, I filtered the data to only include countries with more than 500,000 tons of harvest, made a bar plot, and flipped it sideways so all the names could be read.

```{r fig.align="center", echo = FALSE,fig.width= 25}
big_fisheries %>% 
  ggplot(aes(x = reorder(country , total), y = total , fill = capture)) +
  geom_bar(stat = "identity")+
  coord_flip()+
  labs(title = "Countries Fish Production" , subtitle = "For countries with more than 500,000 tons of harvest" , x = "Country")
```

### For my second attempted improvement I also did a log transformation of tons harvested.

```{r fig.align="center", echo = FALSE,fig.width= 15}
p <- ggplot(big_fisheries , aes(x = reorder(country , total), y = total , fill = capture))
p <- p + scale_y_log10()
p + geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Countries Fish Production" , subtitle = "For countries with more than 500,000 tons of harvest" , x = "Country" , y = "Tons Harvested, log transformed")
```

### I dont really think either of these visualizations are that good. I think to make a better visualization I would need to think about what info I was trying to convey. I think a more interesting visualization would be a graph showing the breakdown of aquaculture vs capture in the 10 countries with the greatest fish harvest.

```{r}
big_fisheries %>% 
  arrange(desc(total))

biggest_fisheries_total <- fisheries %>% 
  filter(total > 3800000)

biggest_fisheries_total %>% 
  arrange(desc(total))
```

```{r}
biggest_fisheries_total_long <- biggest_fisheries_total %>%
  pivot_longer(cols = -country, names_to = "harvest") %>%
  filter(harvest %in% c("capture" , "aquaculture")) %>% 
  mutate(value = as.numeric(value))

biggest_fisheries_total_long %>% 
  arrange(desc(value))
```



```{r}

biggest_fisheries_total_long %>% 
  ggplot(aes(x = reorder(country , -value) , y = value , fill = harvest)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  labs(title = "Tons of Fish Harvested Through Capture or Aquaculture" , subtitle = "Data from the Top Ten Fish Producing Countries" , y = "Tons of Fish Harvested per year" , x = "Country")

```

### Exercise 3

```{r}
library(mosaicData)
data(Whickham)
```

1. These data come from an observational experiment.
2. There are 1314 observations. Each represents a woman.
3. There are 3 variables. Outcome status, smoker status, age.
4. I would expect smoking to have a negative relationship with health outcome.

```{r}
Whickham %>%
  count(smoker, outcome)

Whickham %>% 
  ggplot(aes(x = outcome)) +
  geom_bar() +
  facet_wrap(~smoker)+
  labs(title = "Health Outcomes (Alive or Dead) of Smokers and Non-Smokers at 20 Year Follow Up" , y = "Number of People")

Whickham %>% 
  ggplot(aes(x = smoker , fill = outcome)) +
  geom_bar()+
  labs(title = "Health Outcomes (Alive or Dead) of Smokers and Non-Smokers at 20 Year Follow Up" , y = "Number of People")
```

This graph and probability table shows that people who smoke are more likely to live for another 20 years.

```{r}
Whickham %>% 
  count(smoker , outcome) %>% 
  group_by(smoker) %>% 
  mutate(prop = n / sum(n))

```



```{r}
Whickham <- Whickham %>% 
  mutate(age_cat = case_when(
    age <= 44 ~"18-44" ,
    age > 44 & age <= 64 ~ "44-64" ,
    age > 64 ~ "65+"
  ))

```


```{r}
Whickham %>% 
  ggplot(aes(x = smoker , fill = outcome)) +
  geom_bar()+
  labs(title = "Health Outcomes (Alive or Dead) of Smokers and Non-Smokers at 20 Year Follow Up" , subtitle = "Across 3 Age Ranges" , y = "Number of People")+
  facet_wrap(~age_cat)
```


```{r}
Whickham %>%
  count(smoker, age_cat, outcome)%>% 
  group_by(smoker) %>% 
  mutate(prop = n / sum(n)) 
```


